---
title: "CBA Salary Allocation Dataset Script"
output: html_notebook
---

```{r}
library(tidyverse)
library(Lahman)
setwd("~/Documents/Harvard CSCI E-5b/cba-salary-allocation")
```


Baseball reference WAR data can be found here. Will need to include the baseball reference ID to do the proper linkage
https://www.baseball-reference.com/about/
https://www.baseball-reference.com/data/war_daily_bat.txt
https://www.baseball-reference.com/data/war_daily_pitch.txt

Note: WAR = WAR_rep + WAA
Note: The baseball reference data also contains salary information. How close are the Lahman and Baseball Reference data to each other?

```{r Load WAR Data}
war_pitch <- read_csv("war_daily_pitch.txt")
war_bat <- read_csv("war_daily_bat.txt")
cba <- read_csv("cba_features97-18.csv")
```


```{r Combine Pitch Bat WAR}
pitch <- war_pitch %>%
  select(player_ID, year_ID, WAA, WAR_rep, G) %>%
  mutate(WAR = as.numeric(WAR_rep) + as.numeric(WAA))

bat <- war_bat %>% filter(pitcher == "N") %>%
  select(player_ID, year_ID, WAA, WAR_rep, G) %>%
  mutate(WAR = as.numeric(WAR_rep) + as.numeric(WAA)) %>%
  mutate(WAR_rep = as.numeric(WAR_rep))
  
comb_war <- bind_rows(bat, pitch) %>% group_by(player_ID, year_ID) %>% top_n(1, G)

```

CPI Data from the Federal Reserve because salaries assumed to be nominal. Believe the download only went back to 1985, but can go back further.
https://fred.stlouisfed.org/series/CPIAUCSL
Note: Baseball reference salary data does convert using the CPI  - "Salaries for mid-season call-ups or traded players may not be shown. Salaries converted to alternate years using Consumer Price Index (CPI) data courtesy of the Minneapolis Federal Reserve Board.""

```{r CPI for Real Salary}
CPI <- read_csv("CPIAUCSL.csv")

CPI_year <- CPI %>%  mutate(year = format(DATE, "%Y")) %>% 
  mutate_if(is.character, as.integer) %>%
  group_by(year) %>% summarize(CPI_yr = mean(CPIAUCSL/100)) 
```

Next step will be to get data on when players become Free Agents. We will use the inforamtion available on free agent signing from Baseball Reference. All subsequent years after a player has appeared on the free agency signing list, their status will be encoded as "Veteran". The previous 3 Years will be encoded for "Arbitration" and their first three years will be encoded as "Rookie". A player only gets the "Veteran" status the season for which they signed their free agent contracts. So a player on the 2016 free agency list will have the "Veteran" status starting in 2017.
https://www.baseball-reference.com/leagues/MLB/2019-free-agents.shtml


```{r Player Status}
# https://www.baseball-reference.com/leagues/MLB/2015-free-agents.shtml
fa.2015 <- read.delim(pipe("pbpaste"), sep = ",", header=TRUE)
fa.2015$yearID <- 2015

# https://www.baseball-reference.com/leagues/MLB/2014-free-agents.shtml
fa.2014 <- read.delim(pipe("pbpaste"), sep = ",", header=TRUE)
fa.2014$yearID <- 2014

# https://www.baseball-reference.com/leagues/MLB/2013-free-agents.shtml
fa.2013 <- read.delim(pipe("pbpaste"), sep = ",", header=TRUE)
fa.2013$yearID <- 2013

# https://www.baseball-reference.com/leagues/MLB/2012-free-agents.shtml
fa.2012 <- read.delim(pipe("pbpaste"), sep = ",", header=TRUE)
fa.2012$yearID <- 2012

# https://www.baseball-reference.com/leagues/MLB/2011-free-agents.shtml
fa.2011 <- read.delim(pipe("pbpaste"), sep = ",", header=TRUE)
fa.2011$yearID <- 2011

# https://www.baseball-reference.com/leagues/MLB/2010-free-agents.shtml
fa.2010 <- read.delim(pipe("pbpaste"), sep = ",", header=TRUE)
fa.2010$yearID <- 2010

# https://www.baseball-reference.com/leagues/MLB/2009-free-agents.shtml
fa.2009 <- read.delim(pipe("pbpaste"), sep = ",", header=TRUE)
fa.2009$yearID <- 2009

# https://www.baseball-reference.com/leagues/MLB/2008-free-agents.shtml
fa.2008 <- read.delim(pipe("pbpaste"), sep = ",", header=TRUE)
fa.2008$yearID <- 2008

# https://www.baseball-reference.com/leagues/MLB/2007-free-agents.shtml
fa.2007 <- read.delim(pipe("pbpaste"), sep = ",", header=TRUE)
fa.2007$yearID <- 2007

# https://www.baseball-reference.com/leagues/MLB/2006-free-agents.shtml
fa.2006 <- read.delim(pipe("pbpaste"), sep = ",", header=TRUE)
fa.2006$yearID <- 2006

# https://www.baseball-reference.com/leagues/MLB/2005-free-agents.shtml
fa.2005 <- read.delim(pipe("pbpaste"), sep = ",", header=TRUE)
fa.2005$yearID <- 2005

# https://www.baseball-reference.com/leagues/MLB/2004-free-agents.shtml
fa.2004 <- read.delim(pipe("pbpaste"), sep = ",", header=TRUE)
fa.2004$yearID <- 2004

# https://www.baseball-reference.com/leagues/MLB/2003-free-agents.shtml
fa.2003 <- read.delim(pipe("pbpaste"), sep = ",", header=TRUE)
fa.2003$yearID <- 2003

# https://www.baseball-reference.com/leagues/MLB/2002-free-agents.shtml
fa.2002 <- read.delim(pipe("pbpaste"), sep = ",", header=TRUE)
fa.2002$yearID <- 2002

# https://www.baseball-reference.com/leagues/MLB/2001-free-agents.shtml
fa.2001 <- read.delim(pipe("pbpaste"), sep = ",", header=TRUE)
fa.2001$yearID <- 2001

# https://www.baseball-reference.com/leagues/MLB/2000-free-agents.shtml
fa.2000 <- read.delim(pipe("pbpaste"), sep = ",", header=TRUE)
fa.2000$yearID <- 2000

# https://www.baseball-reference.com/leagues/MLB/1999-free-agents.shtml
fa.1999 <- read.delim(pipe("pbpaste"), sep = ",", header=TRUE)
fa.1999$yearID <- 1999

# https://www.baseball-reference.com/leagues/MLB/1998-free-agents.shtml
fa.1998 <- read.delim(pipe("pbpaste"), sep = ",", header=TRUE)
fa.1998$yearID <- 1998

# https://www.baseball-reference.com/leagues/MLB/1997-free-agents.shtml
fa.1997 <- read.delim(pipe("pbpaste"), sep = ",", header=TRUE)
fa.1997$yearID <- 1997

# https://www.baseball-reference.com/leagues/MLB/1996-free-agents.shtml
fa.1996 <- read.delim(pipe("pbpaste"), sep = ",", header=TRUE)
fa.1996$yearID <- 1996

# https://www.baseball-reference.com/leagues/MLB/1995-free-agents.shtml
fa.1995 <- read.delim(pipe("pbpaste"), sep = ",", header=TRUE)
fa.1995$yearID <- 1995

# https://www.baseball-reference.com/leagues/MLB/1994-free-agents.shtml
fa.1994 <- read.delim(pipe("pbpaste"), sep = ",", header=TRUE)
fa.1994$yearID <- 1994

```

```{r}

fa.all <- rbind(fa.2015, fa.2014, fa.2013, fa.2012, fa.2011, fa.2010, fa.2009,
                fa.2008, fa.2007, fa.2006, fa.2005, fa.2004, fa.2003, fa.2002,
                fa.2001, fa.2000, fa.1999, fa.1998, fa.1997, fa.1996, fa.1995,
                fa.1994) %>% separate(Name, c("name", "bbrefID"), "\\\\") # Escape string headaches https://xkcd.com/1638/
write.csv(fa.all,"faSigning1994-2015.csv")

```
```{r If you already Have Free Agents File}
fa.all <- read_csv("faSigning1994-2015.csv")
```

```{r function to encode veteran status}
vet <- function(data){
  data$veteran <- 0
  
  for(i in unique(data$playerID)){
    c <- 0
    for(j in data[which(data$playerID == i),]$yearID){
      data[which(data$playerID == i & data$yearID == j),]$veteran <- ifelse(c < 7, 0, 1)
      c <- c+1
      #print(c(i, j, c, data[which(data$playerID == i & data$yearID == j),]$veteran))
    }
  }
return(data)
}
```

```{r Create Dataset}
dataset <- Salaries %>%
  filter(salary > 0) %>% # Exclude zero salary so we can use log
  left_join(Master, by = "playerID") %>%
  mutate(Age = yearID - birthYear) %>%
  filter(yearID <= as.numeric(str_sub(finalGame, 1, 4))) %>% # Exclude errors in salary table
  #left_join(Fielding, by = c("yearID", "playerID")) %>%
  left_join(comb_war, by = c("bbrefID" = "player_ID", "yearID" = "year_ID")) %>%
  left_join(CPI_year, by = c("yearID" = "year")) %>%
  left_join(cba, by = yearID) %>%
  mutate(realSalary = salary/CPI_yr) %>% 
  select(yearID, teamID, lgID, playerID, salary, birthYear, birthCountry, nameFirst, nameLast, Age, WAA, WAR_rep, WAR, realSalary, bbrefID) %>%
  vet()
  
```

```{r}
dataset2 <- dataset %>%
  left_join(fa.all, by = c("bbrefID", "yearID")) %>% 
  select(yearID:Yrs)
```

```{r}
dataset <- dataset %>%
  filter(yearID > 1994)
```



```{r}
lmsal <- lm(log(realSalary) ~ WAR + veteran, data = dataset)
summary(lmsal)
```

```{r}
lmsal <- lm(realSalary ~ WAR + veteran, data = dataset)
summary(lmsal)
```


```{r}
test <- dataset %>% 
  select(Age, WAA, WAR, realSalary, veteran, feat1, feat2, feat3, feat4, feat5, feat6, feat7) %>% 
  mutate(WAA = as.numeric(WAA)) %>%
  cor() %>% round(3) %>% 
  as.data.frame() %>% 
  add_rownames("var1") %>%
  gather("var2", "cor", 1:13) %>%
  ggplot(aes(var1, var2, fill = cor)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red") +
  geom_text(aes(var1, var2, label = cor), color = "black", size = 3)
```


```{r}
dataset %>% 
  select(Age, WAR, realSalary, veteran, feat1, feat2, feat3, feat4, feat5, feat6, feat7) %>% 
  sapply(as.numeric) %>%
  #mutate(WAR= as.integer(WAR)) %>%
  cor() %>% round(3)%>% 
  as.data.frame() %>% 
  add_rownames("var1") %>%
  gather("var2", "cor", 2:12) %>%
  ggplot(aes(var1, var2, fill = cor)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red") +
  geom_text(aes(var1, var2, label = cor), color = "black", size = 3)
```



```{r}
dataset %>% 
  select(Age, WAR, realSalary, veteran) %>% 
  sapply(as.numeric) %>%
  #mutate(WAR= as.integer(WAR)) %>%
  cor() %>% round(3)%>% 
  as.data.frame() %>% 
  add_rownames("var1") %>%
  gather("var2", "cor", 2:5) %>%
  ggplot(aes(var1, var2, fill = cor)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red") +
  geom_text(aes(var1, var2, label = cor), color = "black", size = 3)
```


```{r}

```






