---
title: "MLB Salaries"
output:
  word_document: default
  html_document: default
---

```{r setup, include=TRUE}
library(tidyverse)
library(Lahman)
library(scales)
```


After looking at the trends in hitting from class 3 and the Manny Machado $300 million contract, I was curious about trends in MLB salaries over the course of MLB history. The Lahman database includes a table with this information going back to 1985. Lahman says there is no real coverage of baseball salaries until this year. This is unfortunate, because I won't be able to see how new collective bargaining regimes might have impacted it (ie. Free Agency around 1976). May see some impact from the 1994 stike. 

I believe this data is in nominal terms and may need to be rescaled to real salaries using the CPI. https://fred.stlouisfed.org/series/CPIAUCSL 

```{r}
summary(Salaries)
```

## Average salary by year - 1985-2018
```{r}
pay <-  Salaries %>% group_by(yearID) %>% summarize(avgPay = mean(salary))

ggplot(pay, aes(yearID, avgPay)) + 
  geom_point() + 
  ggtitle("Average Nominal Salary By Year")
```

Its not specified in the Lahman database whether the salary data is in real or nominal value. Given the spread, its highly likely that the data is nominal. So, I took CPI data from the FRED economic database to convert the salary data to nominal values. 

```{r}
CPI <- read_csv("CPIAUCSL.csv")

CPI_year <- CPI %>%  mutate(year = format(DATE, "%Y")) %>% 
  mutate_if(is.character, as.integer) %>%
  group_by(year) %>% summarize(CPI_yr = mean(CPIAUCSL/100)) 

pay <- left_join(pay, CPI_year, by = c("yearID" = "year")) %>% mutate(realSalary = avgPay / CPI_yr)

ggplot(pay, aes(yearID, realSalary)) + 
  geom_point() + 
  ggtitle("MLB Players Real Salary - (CPI 1985=100)")

```




```{r}
ggplot(Salaries, aes(yearID, salary)) +
  geom_smooth() +
  geom_point() + 
  scale_y_continuous(labels = dollar) +
  ggtitle("Plot of Nominal Salaries with Conditional Mean")
```

```{r}
ggplot(Salaries, aes(x=factor(yearID), y=salary)) +
  geom_boxplot() +
  xlab("Year") +
  ylab("Annual Salary") +
  scale_x_discrete(breaks=unique(Salaries$yearID)[c(TRUE, FALSE)]) +
  scale_y_continuous(labels = dollar) + 
  ggtitle("Nominal Salary with Boxplots")
```


```{r}

Salaries %>% 
  left_join(Fielding, by = c("yearID", "playerID")) %>%
  group_by(yearID, POS) %>% 
  summarize(avgPay = mean(salary)) %>% 
  ggplot(aes(x=yearID, y=avgPay, col=POS)) + 
  geom_point() + 
  geom_line() +
  xlab("Year") +
  ylab("Annual Salary") +
  scale_x_continuous() +
  scale_y_continuous(labels = dollar) + 
  ggtitle("Average Salary By Position and Year")
```

```{r}
Salaries %>%
  left_join(Master, by = "playerID") %>%
  filter(yearID >= 2005) %>%
  mutate(Age = yearID - birthYear) %>%
  filter(yearID <= as.numeric(str_sub(finalGame, 1, 4))) %>%
  group_by(Age, yearID) %>%
  summarize(avgPay = mean(salary)) %>%
  ggplot(aes(x=Age, y=avgPay, col=yearID, group=yearID)) + 
  #geom_point() + 
  geom_line() +
  xlab("Age") +
  ylab("Annual Salary") +
  scale_x_continuous() +
  scale_y_continuous(labels = dollar) + 
  ggtitle("Average Salary By Age")

```

```{r}
Salaries %>%
  left_join(Master, by = "playerID") %>%
  filter(yearID == 1990 | yearID == 2000 | yearID == 2010 | yearID == 2015) %>%
  mutate(Age = yearID - birthYear) %>%
  filter(yearID <= as.numeric(str_sub(finalGame, 1, 4))) %>%
 mutate(yearID = as.character(yearID)) %>%
  group_by(Age, yearID) %>%
  add_count(Age) %>%
  ggplot(aes(x=Age, y=n, col=yearID, group=yearID)) + 
  #geom_point() + 
  #geom_line() +
  geom_smooth() +
  xlab("Age") +
  ylab("Total Number of Players") +
  scale_x_continuous() +
  scale_y_continuous() + 
  ggtitle("MLB Players by Age")
```

```{r}
test <- Salaries %>%
  left_join(Master, by = "playerID") %>%
  filter(yearID >= 2010) %>%
  mutate(Age = yearID - birthYear) %>%
  filter(yearID <= as.numeric(str_sub(finalGame, 1, 4))) %>%
  group_by(Age, yearID) %>%
  add_count(Age)
```

